Breareal

### 2022-08-05 indicators/breareal.Rmd
rm chunk names

```{r setup, include=FALSE, message=FALSE}
library(knitr)
library(sf)
library(raster) # only needed for creating boundary boxes, which can be done using sf
library(dplyr)
library(lwgeom)
library(tmap)
library(dplyr)
#library(ggplot2)
#library(data.table)
knitr::opts_chunk$set(echo = TRUE)
```

@@

#### Glacial atlas
This is the new glacial atlas from 2018 and 2019 based on Sentinell (ref. Liss Marie Andreassen, NVE)
```{r import1}
# This first part is if you are running this from the NINA R server. 
#breatlas <- sf::st_read('/data/P-Prosjekter/41201042_okologisk_tilstand_fastlandsnorge_2020_dataanaly/fjell2021/data/breatlas/breatlas_2018_2019/Breatlas_20182019_temp20210922_lma.shp')

@@

There are 6915 polygons. Let's calculate the area of the polygons and study the distribution of size classes. 
```{r breareal}
breatlas$area <- st_area(breatlas)
par(mfrow=c(1,2))
hist(breatlas$area)
plot(breatlas$area)
```

@@

#### Map of Norway
This map is stored on the GitHub repository under _indicators/data/_. 
```{r import2}
nor <- sf::read_sf("data/outlineOfNorway_EPSG25833.shp")

@@

```{r import3}
#n50 <- sf::st_read('/data/P-Prosjekter/41201042_okologisk_tilstand_fastlandsnorge_2020_dataanaly/fjell2021/data/breatlas/n50/cryoclim_GAO_NO_1952_1985_UTM_33N.shp')%>%
#  st_transform(crs = crs(breatlas))

@@

```{r plotter-alt}

@@

```{r import4}
reg <- st_read("data/regions.shp")
st_crs(reg)
```

@@

There is a convinient function to fix this.
```{r make-valid}
# requires lwgeom
breatlas2 <- st_make_valid(breatlas)
table(st_is_valid(breatlas2))

@@

Here is one method for getting the glacial area for each region.
```{r intersection}
# This intersection operation is a bit heavy, so I do it once and store it locally so that I can import it again quickly. 
#brealtas_reg <- st_intersection(breatlas, reg)
#saveRDS(brealtas_reg, "indicators/data/brealtas_reg_helperfile.rds")

@@

Calculating the area of the polygons
```{r plotting-area-per-region}
#Calculating the area of the polygons
brealtas_reg$area_crop <- st_area(brealtas_reg)

@@

#### Same for N50
```{r intersection2}
n50x <- st_make_valid(n50)
#n50_reg <- st_intersection(n50x, reg)
#saveRDS(n50_reg, "indicators/data/n50_helperfile.rds")
n50_reg <- readRDS("data/n50_helperfile.rds")
```

@@

Thats good. 
Now I will plot the reference values (n50) and the contemporary data together.
```{r plotting-ref-and-ind}
n50_reg$area_crop <- st_area(n50_reg)

@@

### Scaled indicator values
Making a table with the indicator values.
```{r skalert}
myTbl <- tibble(region           = names(bretabell),
                indikator        = bretabell,
                referanseverdi   = n50tabell)
myTbl$skalert_indikator <- myTbl$indikator/myTbl$referanseverdi

>>>>

```{r, include=FALSE, message=FALSE}
library(knitr)
library(sf)
library(raster) # only needed for creating boundary boxes, which can be done using sf
library(dplyr)
library(lwgeom)
library(tmap)
library(dplyr)
#library(ggplot2)
#library(data.table)
knitr::opts_chunk$set(echo = TRUE)
```

@@

#### Glacial atlas
This is the new glacial atlas from 2018 and 2019 based on Sentinell (ref. Liss Marie Andreassen, NVE)
```{r}
# This first part is if you are running this from the NINA R server. 
#breatlas <- sf::st_read('/data/P-Prosjekter/41201042_okologisk_tilstand_fastlandsnorge_2020_dataanaly/fjell2021/data/breatlas/breatlas_2018_2019/Breatlas_20182019_temp20210922_lma.shp')

@@

There are 6915 polygons. Let's calculate the area of the polygons and study the distribution of size classes. 
```{r}
breatlas$area <- st_area(breatlas)
par(mfrow=c(1,2))
hist(breatlas$area)
plot(breatlas$area)
```

@@

#### Map of Norway
This map is stored on the GitHub repository under _indicators/data/_. 
```{r}
nor <- sf::read_sf("data/outlineOfNorway_EPSG25833.shp")

@@

```{r}
#n50 <- sf::st_read('/data/P-Prosjekter/41201042_okologisk_tilstand_fastlandsnorge_2020_dataanaly/fjell2021/data/breatlas/n50/cryoclim_GAO_NO_1952_1985_UTM_33N.shp')%>%
#  st_transform(crs = crs(breatlas))

@@

```{r}

@@

```{r}
reg <- st_read("data/regions.shp")
st_crs(reg)
```

@@

There is a convinient function to fix this.
```{r}
# requires lwgeom
breatlas2 <- st_make_valid(breatlas)
table(st_is_valid(breatlas2))

@@

Here is one method for getting the glacial area for each region.
```{r}
# This intersection operation is a bit heavy, so I do it once and store it locally so that I can import it again quickly. 
#brealtas_reg <- st_intersection(breatlas, reg)
#saveRDS(brealtas_reg, "indicators/data/brealtas_reg_helperfile.rds")

@@

Calculating the area of the polygons
```{r}
#Calculating the area of the polygons
brealtas_reg$area_crop <- st_area(brealtas_reg)

@@

#### Same for N50
```{r}
n50x <- st_make_valid(n50)
#n50_reg <- st_intersection(n50x, reg)
#saveRDS(n50_reg, "indicators/data/n50_helperfile.rds")
n50_reg <- readRDS("data/n50_helperfile.rds")
```

@@

Thats good. 
Now I will plot the reference values (n50) and the contemporary data together.
```{r}
n50_reg$area_crop <- st_area(n50_reg)

@@

### Scaled indicator values
Making a table with the indicator values.
```{r}
myTbl <- tibble(region           = names(bretabell),
                indikator        = bretabell,
                referanseverdi   = n50tabell)
myTbl$skalert_indikator <- myTbl$indikator/myTbl$referanseverdi
